%%
%% GMU LaTeX PhD Dissertation Format Template
%%
%% Developed by:
%%      Daniel O. Awduche and Christopher A. St. Jean
%%      Communications and Networking Lab
%%      Dept. of Electrical and Computer Engineering
%%
%% Notes on usage can be found in the accompanying USAGE_NOTES.txt file.
%%
%%**********************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either
%% expressed or implied; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall any contributor to this code be liable for any damages
%% or losses, including, but not limited to, incidental, consequential, or
%% any other damages, resulting from the use or misuse of any information
%% contained here.
%%**********************************************************************
%%
%% $Id: GMU_dissertation_template.tex,v 1.7 2007/05/02 02:20:38 Owner Exp $
%%

\documentclass[11 pt]{report}

%%  The file ``gmudissertation.sty''  is the GMU latex style file and
%%   should be placed in the same directory as your LaTeX files
\usepackage{gmudissertation}

%%
%% other packages that need to be loaded
%%
\usepackage{graphicx}                    %   for imported graphics
\usepackage{amsmath}                     %%
\usepackage{amsfonts}                    %%  for AMS mathematics
\usepackage{amssymb}                     %%
\usepackage{amsthm}                      %%
\usepackage[normalem]{ulem}              %   a nice standard underline package
\usepackage[noadjust,verbose,sort]{cite} %   arranges reference citations neatly
\usepackage{setspace}                    %   for line spacing commands
\usepackage{subfigure}						 %   FIGURE PLACEMENT
\usepackage{natbib}
\usepackage{microtype}

%% The file ``mydissertationabbrev.sty'' is an (optional) personalized file that
%% may contain any and all LaTeX command (re)definitions that will be used
%% throughout the document
%\usepackage{mydissertationabbrev}

\beforedoc

\begin{document}

%% In this section, all of the user-specific fields to be used in the
%% title pages are set
\title{Validation of \\ Magnetospheric Magnetohydrodynamic Models}
\onelinetitle{Validation of Magnetospheric Magnetohydrodynamic Models}
\author{Brian Curtis}
\degree{Doctor of Philosophy}
\doctype{Dissertation}
\dept{School of Physics, Astronomy, and Computational Sciences}
\discipline{Computational Science and Informatics}

\seconddeg{Master of Science}
\seconddegschool{George Mason University}
\seconddegyear{2010}

\firstdeg{Bachelor of Science}
\firstdegschool{State University of New York at Oswego}
\firstdegyear{2007}

\degreeyear{2014}

% Note: semester name should be written in its full-form. For example, Fall Semester, not just Fall.
\degreesemester{Spring Semester}

\advisor{Dr. Robert Weigel}

\firstmember{Dr. Jie Zhang}

\secondmember{Dr. Arthur Poland}

\thirdmember{Dr. Ruixin Yang}

\depthead{Dr. Chi Yang}

\assocdean{Dr. Donna M. Fox}

\deanITE{Dr. Peggy Agouris}

%%
%% Introductory pages
%%

% Note: The signature sheet is set according to the requirements of the Volgenau School of
% Information Technology and Engineering. If your college/school requirement is different,
% please make appropriate changes in the "signaturepage" section of gmudissertation.sty file.
\signaturepage

\titlepage

% copyright technically optional but should be included in to avoid potential pagination problems
\copyrightpage

%%
%% Dedication page
%%

\dedicationpage

\noindent I dedicate this dissertation to my friends and family who have helped
push me along this journey and made it worthwhile. Specifically my wife,
Tamarah, who has dealt with my hectic work life and loved and supported me so
much, making sure that home was truly a comforting and restoring place for my
mind and body. Also, my parents Robert and Deborah Curtis and brother Jeff
Curtis, who have supported me throughout all my years in education and shown me
the importance every little detail makes in the big picture. Finally, my
expected daughter, Aurora, who gives me inspiration to become as great of a
human being as I can, to guarantee she has the best role model anyone could ask
for.

%%
%% Acknowledgements
%%

\acknowledgementspage

\noindent I would like to thank my dissertation committee: Dr. Robert Weigel,
Dr. Art Poland, Dr. Jie Zhang, and Dr. Ruixin Yang for agreeing to give me their
time and effort. Special thanks to Dr. Robert Weigel who has made this whole
graduate experience possible, whose guidance and insight has prepared
me well for scientific pursuit and what to expect after graduation.

%%
%% Table of contents, list of tables, and lists of figures
%%

\tableofcontents

\listoftables

\listoffigures

% % % Abstract %
\abstractpage

Magnetospheric magnetohydrodynamic (MHD) models are commonly used for both
prediction and modeling of Earth's magnetosphere.  To date, very little
validation has been performed to determine their limits, uncertainties, and
differences.  In this work,
we performed a comprehensive analysis using several commonly used validation
techniques in the atmospheric sciences to MHD-based models of Earth's
magnetosphere for the first time. The validation techniques of \texit{parameter
variability/sensitivity} analysis and \texit{comparison to other models} were used on the OpenGGCM, BATS-R-US, and SWMF
magnetospheric MHD models to answer several questions about how these
models compare.  The questions include: (1) the difference between the model's
predictions prior to and following to a reversal of $B_z$ in the upstream
interplanetary field (IMF) from positive to negative, (2) the influence of the
preconditioning duration, and (3) the differences between models under extreme
solar wind conditions. A differencing visualization tool was developed and used to
address these three questions.

\abstractmultiplepage We find: (1) For a reversal in $B_z^{IMF}$ from positive
to negative, the OpenGGCM magnetopause is closest to Earth as it has the weakest
magnetic pressure near-Earth. The differences in magnetopause positions between
BATS-R-US and SWMF are explained by the influence of the ring current, which is
included in SWMF. Densities are highest for SWMF and lowest for OpenGGCM. The
OpenGGCM tail currents differ significantly from BATS-R-US and SWMF; (2) A
longer preconditioning time allowed the magnetosphere to relax more, giving
different positions for the magnetopause with all three models before the
$B_z^{IMF}$ reversal. There were differences greater than 100\% for all three
models before the $B_z^{IMF}$ reversal. The differences in the current sheet
region for the OpenGGCM were small after the $B_z^{IMF}$ reversal. The BATS-R-US
and SWMF differences decreased after the $B_z^{IMF}$ reversal to near zero; (3)
For extreme conditions in the solar wind, the OpenGGCM has a large region of
Earthward flow velocity ($U_x$) in the current sheet region that grows as time
progresses in a compressed environment. BATS-R-US $B_z$, $\rho$ and $U_x$ stabilize to a near constant value approximately one hour into the run under high compression conditions. Under
high compression, the SWMF parameters begin to oscillate approximately 100 minutes into
the run. All three models have similar magnetopause positions under low pressure
conditions. The OpenGGCM current sheet velocities along the Sun-Earth line are
largest under low pressure conditions.

The results of this analysis indicate the need for accounting for model
uncertainties and differences when comparing model predictions with data, provide error bars on
model prediction in various magnetospheric regions, and show that the
magnetotail is sensitive to the preconditioning time.

%% Be sure to leave a line of whitespace immediately before this line!!!!!
%% (If this comment segment runs together with the preceeding text, you might
%%  see the second page of the abstract numbered "0".)
%%
%% If the abstract is more than one page, then place this line PRECISELY
%% at the page break; otherwise, comment it out.  (See note about this line
%% in the usage notes.)
%%


%%
%%  the main body of the dissertation
%%
\startofchapters

%% include the chapters one by one (or paste the chapter text in directly if desired)
\include{Introduction}
\include{BzFlip}
\include{Preconditioning}
\include{ExtremeConditions}
%%\include{Research}
%%\include{Results}
%%\include{Conclusion}

%% Note: appendix is now put before bibliography.
%% include the following directives if there are any appendices
\appendix
\appendixeqnumbering
\include{Appendix}

%%
%%  bibliography
%%

%% list all of the BibTeX files here for the WinEdt project (if applicable)
%GATHER{bibfile.bib}

%% any bibliography style can be used, but IEEEtran.bst is ideally suited to
%% electrical engineering references

\bibliographystyle{agufull}
%%\bibliographystyle{IEEEtran}
%%\bibliography{IEEEfull,bibfile}
\bibliography{bibfile}

%%
%% curriculum vitae
%%
\cvpage

\noindent Brian Curtis was born in 1985 in Rochester, New York. He
graduated from the State University of New York at Oswego with a Bachelor's
degree in Meteorology in 2007. He also received a master's degree in
Computational Sciences from George Mason University in Fairfax, Virginia in
2010. He was employed for 7 years as a research assistant under the instruction
of Dr. Robert Weigel. His research involves the validation of
magnetospheric magnetohydrodynamic space weather models. During his time at
George Mason University, he worked on scientific outreach through an annual space
weather forecasting contest, was a summer research assistant at the space
weather prediction center in Boulder, Colorado in 2009, and in the summer of 2013
was a science collaborator at the National Aeronautics and Space Administration Goddard Space
Flight Center in Greenbelt, Maryland.

In addition Brian has attended the:\\
- Geospace Environment Modeling summer conference in Snowmass, CO in 2009.\\
- Center for Integrated Space weather Modeling summer school in 2008.\\
- American Meteorological Society annual conference from 2004 to 2007.
\end{document}
